---
title: "Re-analysis"
format: html
editor: visual
---

# Re-analysis of paper \*Mediated phonological–semantic priming in spoken word production: Evidence for cascaded processing from picture–word interference\*

# Experiment 2

## Preamble: Loading packages and configuration

```{r}
#| label: data_and_libraries
#| echo: false

# just run this code chunk
# function to ignoring the setting of the relative path below when knitting
run_if_not_knitting <- function(expr) {
  if (!isTRUE(getOption("knitr.in.progress"))) {
    eval(expr)
  }
}

# nifty code using the pacman package
# it checks if the packages specified below are installed, if not, they will be installed, if yes, they will be loaded
if (!require("pacman")) install.packages("pacman")
pacman::p_load(rstudioapi, tidyverse, lme4, trimr)

# set the current working directory to the one where this file is
run_if_not_knitting(current_working_dir <- dirname(rstudioapi::getActiveDocumentContext()$path))
run_if_not_knitting(setwd(current_working_dir))

```

# Experiment 1

## Preparation

downloading data_written from their osf repository: https://osf.io/65f7w/overview and storing in the same folder as this quarto document

```{r preparing data}
# loading the data file
aud_data_raw <- read_csv("data_auditory.csv")

# filtering data following their description in paper
aud_data_filtered <- aud_data_raw|>
  # excluding speech errors
  filter(accuracy == 1) |>
  # excluding empty audio files
  filter(rt > 0)

# checking how many trials got excluded
count(aud_data_raw)-count(aud_data_filtered)
# We excluded 67 trials just as they said in the paper

# continuing filtering the data with the "trimr" package
aud_data_trimmed <- aud_data_filtered |>
  # excluding RT below 250ms
  sdTrim(minRT = 250,
         # excluding RT above mean+3*sd per participant
         sd = 3,
         perCondition = FALSE,
         perParticipant = TRUE,
         returnType = "raw")
# in paper they say that the excluced -3sd too,
# but that cannot be done with the sdTrim function

# checking how much data we excluded:
count(aud_data_filtered)-count(aud_data_trimmed)
# while we only excluded 48 trials, the paper reports that 57 trials where removed 

# renaming data to make it easier
data_ex1 <- aud_data_trimmed

```

## Reproducing Table 3

```{r}
# mean RT for each condition

mean_RT <- data_ex1 |>
  # grouping by condition
  group_by(condition)|>
  # calculating mean of reaction time and rounding on 2 digits
  summarize(mean = round(mean(rt),2),
            sd = round(sd(rt),2))
  
  
  
# making a df that displays data just like table 3 in their paper
# reordering the conditions
condition <- c("Categorically related",
               "Unrelated",
               "Semantic inference",
               "Mediated phonological-semantical",
               "Unrelated",
               "Semantic inference")
# extracting mean rt in the same order
M <- c(mean_RT$mean[1],
          mean_RT$mean[3],
          mean_RT$mean[1]-mean_RT$mean[3], 
          mean_RT$mean[2],
          mean_RT$mean[4],
          mean_RT$mean[2]-mean_RT$mean[4])
# extracting sd in the same order
SD <- c(mean_RT$sd[1],
        mean_RT$sd[3],
        "",
        mean_RT$sd[2],
        mean_RT$sd[4],
        "")
# making a tibble out if the objects containing the new order
table_3 <- tibble(Distractor_Type = condition,
       M = M,
       SD = SD)

table_3
# The mean rt for each condition if not the same as reported in the paper, for the Categorically related and unrelated mean rt, the differences are only minor (0.26 and 0.76 ms), for the mediated phonological-semantical related and unrelated distractor types, the differences are a bigger (1.33 and 0.77 ms).
# The results reported for the sd cannot be reproduced at all. Differences between their and our results are around 70ms for all 4 categories.
           

```

## LME analysis

the original data only has a colum "condition" with all for combinations across distractor type and distractor relatedness. In their paper, however, the researchers report to have two fixed effects: distractor type and distractor relatedness Therefore, the data frame needs to be adjusted for the model to be able to know what to do

```{r}
# changing data frame
data_adjusted <- data_ex1 |>
  mutate(distractor type)

# preparing the data for sum-coding



contrasts(data_trimmed$condition) <- 

```

## notes about their analysis:

# Experiment 2

## Preparation

downloading data_written from their osf repository: https://osf.io/65f7w/overview and storing in the same folder as this quarto document

```{r preparing data}
# loading the data file
data_raw <- read_csv("data_written.csv")

# filtering data following their description in paper
data_filtered <- data_raw|>
  # excluding speech errors
  filter(accuracy == 1) |>
  # excluding empty audio files
  filter(rt > 0)

# checking how many trials got excluded
count(data_raw)-count(data_filtered)
# We excluded 108 trials just as they said in the paper

# continuing filtering the data with the "trimr" package
data_trimmed <- data_filtered |>
  # excluding RT below 250ms
  sdTrim(minRT = 250,
         # excluding RT above mean+3*sd per participant
         sd = 3,
         perCondition = FALSE,
         perParticipant = TRUE,
         returnType = "raw")
# in paper they say that the excluced -3sd too,
# but that cannot be done with the sdTrim function

# checking how much data we excluded:
count(data_filtered)-count(data_trimmed)
# same amount of trials as in paper

```

## Reproducing Table 4

```{r}
# mean RT for each condition

mean_RT <- data_trimmed |>
  # grouping by condition
  group_by(condition)|>
  # calculating mean of reaction time and rounding on 2 digits
  summarize(mean = round(mean(rt),2),
            sd = round(sd(rt),2))
  
  
  
# making a df that displays data just like table 4 in their paper
# reordering the conditions
condition <- c("Categorically related",
               "Unrelated",
               "Semantic inference",
               "Mediated phonological-semantical",
               "Unrelated",
               "Semantic inference")
# extracting mean rt in the same order
M <- c(mean_RT$mean[1],
          mean_RT$mean[3],
          mean_RT$mean[1]-mean_RT$mean[3], 
          mean_RT$mean[2],
          mean_RT$mean[4],
          mean_RT$mean[2]-mean_RT$mean[4])
# extracting sd in the same order
SD <- c(mean_RT$sd[1],
        mean_RT$sd[3],
        "",
        mean_RT$sd[2],
        mean_RT$sd[4],
        "")
# making a tibble out if the objects containing the new order
table_4 <- tibble(Distractor_Type = condition,
       M = M,
       SD = SD)

table_4
# The mean rt for each condition if not the same as reported in the paper, for the Categorically related and unrelated mean rt, the differences are only minor (0.15 and 0.25 ms), for the mediated phonological-semantical related and unrelated distractor types, the differences are a bigger (1.14 and 1.2 ms).
# The results reported for the sd cannot be reproduced at all. Differences between their and our results are around 50ms for all 4 categories.
           

```

## LME analysis

the original data only has a colum "condition" with all for combinations across distractor type and distractor relatedness. In their paper, however, the researchers report to have two fixed effects: distractor type and distractor relatedness Therefore, the data frame needs to be adjusted for the model to be able to know what to do

```{r}
# changing data frame
data_adjusted <- data_trimmed |>
  mutate(distractor type)

# preparing the data for sum-coding



contrasts(data_trimmed$condition) <- 

```

## notes about their analysis:
