---
title: "Tang_part5"
author: "Tangerine"
format: html
editor: visual
---

# Buddy challenge part 5: SESOI, Bayes Factor and ROPEs

TASK, DELETE LATER:

Define the smallest effect size of interest for the original research question and do a literature research to justify your choice.

rerun your hypothesis test combining the concepts of Bayes Factor (BF) and Regions of Practical Equivalence (ROPE) (according to Roettger & Franke's method, 2025)

run a sensitivity analyses for different ROPES and priors; write up the results accordingly. save the analysis and report in your Github repository Each buddy group leader additionally submits their report as a markdown (.qmd) and knitted pdf file here.

Naming convention for files: groupname_part5

END OF DELETE LATER

## Seting up working directory, packages and data

```{r, packages and workind directory}

# just run this code chunk
# function to ignoring the setting of the relative path below when knitting
run_if_not_knitting <- function(expr) {
  if (!isTRUE(getOption("knitr.in.progress"))) {
    eval(expr)
  }
}

# nifty code using the pacman package
# it checks if the packages specified below are installed, if not, 
#they will be installed, if yes, they will be loaded
if (!require("pacman")) install.packages("pacman")
pacman::p_load(rstudioapi, tidyverse, bayesrules,
               tidybayes, trimr, brms, bayesplot, bayestestR, modelr)

# set the current working directory to the one where this file is
run_if_not_knitting(current_working_dir <- dirname(rstudioapi::getActiveDocumentContext()$path))
run_if_not_knitting(setwd(current_working_dir))

# options to increase efficiency of brms models (optional)
#options(brms.backend = "cmdstanr")
options(mc.cores = parallel::detectCores())


```

```{r, data loading}
#| label: prep data for task 1
#| echo: false 

# This code is the same as for part 4 since the wrangling is the same
aud_raw <- read_csv("reanalysis/data_auditory.csv")

aud  <-  aud_raw |>
  # excluding speech errors
  filter(accuracy == 1) |>
  # excluding empty audio files
  filter(rt > 0) |> 
  # excluding RT below 250ms
  sdTrim(minRT = 250,
         # excluding RT above mean+3*sd per participant
         sd = 3,
         perCondition = FALSE,
         perParticipant = TRUE,
         returnType = "raw") |>
  # Changing the name of the column "sound" 
  # to match the original naming in Mascelloni et al
  mutate(distractor_word = sound) |>
  # removing the column named sound
  select(-sound) |>
  # creating categories needed for the model
  mutate(distractor_type = case_when(condition == "category" ~ "Categorical",
                                     condition == "unrelated_cat" ~ "Categorical",
                                     condition == "mediated" ~ "Mediated",
                                     condition == "unrelated_med" ~ "Mediated"),
         # creating a column for distractor relatedness
         distractor_relatedness = case_when(condition == "category" ~ "related",
                                 condition == "mediated" ~ "related",
                                 condition == "unrelated_cat" ~ "unrelated",
                                 condition == "unrelated_med" ~ "unrelated")) |>
  # coding new columns as factors
  mutate(distractor_type = as.factor(distractor_type),
         distractor_relatedness = as.factor(distractor_relatedness)) |>
  # removing unnecessary columns for aesthetics
  select(-condition, -run, -SOA, -accuracy)

# double checking levels
levels(aud$distractor_type)
levels(aud$distractor_relatedness)
# changing contrasts for sum-coding
contrasts(aud$distractor_type) <- c(-0.5, 0.5)
contrasts(aud$distractor_relatedness) <- c(0.5, -0.5)

```

## Smallest effect size of interest

We focus our buddy challenge on Experiment 1 from Mascelloni et al. (2021). The original specific research question that we focused on for this assignment was whether a mediated semantic interference effect could be shown using a picture naming methodology. As an example of what this means, it would be participants being slower (taking more milliseconds) at naming the picture of a bat after being exposed to the prime "dog" due to the interference created by the potentially mediated activated word "cat". The overarching big picture research question is whether a cascade model of lexical retrieval is more correct than a sequential model.

Establishing a SESOI has been challenging. In our case, we need to agree on the number of ms that we would find convincing as a difference between conditions, beyond pure chance or measurement error. In order to do so, we looked at the literature on mediated phonological-semantic priming to try to find meta-analysis that would give us an idea of what size we would consider meaningful. We unfortunately found no such meta-analysis. The closest we could find was a book chapter by Kerr et al. (in press). However, said chapter does not provide anything that would give us any help choosing a SESOI. The closest we find to putting things into number is "lexical processing around 175ms and phonological form encoding some 100ms later around 250-300ms" (Kerr et al., in press:19). Given that the range given for the phonological encoding varies by 50 ms, this is potentially the closest we get to something resembling a number of ms to have as a meaningful difference.

We tried to look at the measurement error of the equipment itself to give us a pointer towards what we could consider a SESOI. The program used in Mascelloni et al. (2021) to quantify the reaction times is Chronset (Roux et al., 2017). We were unable to find any reported measurement error of that software.

Therefore, we decided to define our SESOI based on the wiggle room provided in Kerr et al. (2021) for phonological form encoding, which is 50 ms.

## Rerun of our hypothesis test using BF and ROPE

```{r, reruning your hypothesis test combining the concepts of Bayes Factor (BF) and Regions of Practical Equivalence (ROPE)}


# Here we set up the model we used in part 4, with weakly informative priors
model <- bf(rt ~ distractor_type * distractor_relatedness + 
                    (1 + distractor_type * distractor_relatedness | participant) + 
                    (1 + distractor_type * distractor_relatedness | pic) + 
                    (1 + distractor_relatedness | distractor_word))

get_prior(model, aud, family = "gaussian")

weak_priors <- c(prior(normal(718, 220), class = "Intercept"),
                 prior(normal(0, 220), class = "b", coef = "distractor_relatedness1"),
              prior(normal(0, 220), class = "b", coef = "distractor_type1"),
              prior(normal(0, 220), class = "b", coef = "distractor_type1:distractor_relatedness1"),
              prior(lkj(1), class = 'cor'),
              # changing the parameter of the sd-priors from 3 to 300
              prior(cauchy(0, 300), class = 'sd', group = 'distractor_word'),
              prior(cauchy(0, 300), class = 'sd', group = 'participant'),
              prior(cauchy(0, 300), class = 'sd', group = 'pic'),
              prior(normal(0, 138), class = "sigma"))
              
proper_model <- brm(
  formula = model,
  data = aud,
  prior = weak_priors,
  family = gaussian(),
  chains = 4,
  iter = 2000,
  warmup = 1000,
  thin = 1, 
  seed = 42,
  cores = 4        
)
# looking at model output
proper_model
# chaning the sd-priors had a big influence on the 95% confidence interval of the
# intercept, which now is [701.77, 803.42]. It increased arroud 12ms compared to
# the previous model [709.81, 799.08]. However, the mean of the estimate didn't
# change much, only 0.78 ms. For the distractor type, the estimate changes only 
# 0.32ms, the 95% CI changed from [-7.42, 27.92] to [-7.33, 29.52]. The change does
# not lead to differnt signs and is quite small compared to the change foundin the
# other coefficiant, distractor relatedness where it now is [7.95, 32.41] compared
# to the previos 95% CI of [10.67, 30.34]. The interaction is affected the most
where the interval increased from [-27.12, 11.64] to [-30.12, 14.10].


# using our SESOI to define our ROPEs
rope <- c(-50,50)
bf_rope_type <- bayesfactor_rope(posterior = proper_model,
                 null = rope,
                 parameter = 'b_distractor_type1')
bf_rope_type
bf_rope_related <- bayesfactor_rope(posterior = proper_model,
                 null = rope,
                 parameter = 'b_distractor_relatedness1')
bf_rope_related
bf_rope_inter <- bayesfactor_rope(posterior = proper_model,
                 null = rope,
                 parameter = 'b_distractor_type1:distractor_relatedness1')
bf_rope_inter

# even when setting a seed, the outouts of the bayesfactor_rope() function changes 
# Our results where the following numbers:
# type 0.00000807
# relatedness 0.0000578
# interaction 0.0000507
```

## Run a sensitivity analyses for different ROPES and priors

```{r, sensitivity analyses - different priors}

# setting priors that are half as wide as our original priors
priors_110 <- c(prior(normal(718, 220), class = "Intercept"),
                 prior(normal(0, 110), class = "b", coef = "distractor_relatedness1"),
              prior(normal(0, 110), class = "b", coef = "distractor_type1"),
              prior(normal(0, 110), class = "b", coef = "distractor_type1:distractor_relatedness1"),
              prior(lkj(1), class = 'cor'),
              # changing the parameter of the sd-priors from 3 to 300
              prior(cauchy(0, 300), class = 'sd', group = 'distractor_word'),
              prior(cauchy(0, 300), class = 'sd', group = 'participant'),
              prior(cauchy(0, 300), class = 'sd', group = 'pic'),
              prior(normal(0, 138), class = "sigma"))
# running model with new priors
narrow_model <- brm(
  formula = model,
  data = aud,
  prior = priors_110,
  family = gaussian(),
  chains = 4,
  iter = 2000,
  warmup = 1000,
  thin = 1, 
  seed = 42,
  cores = 4        
)

# re-running our BF with same rope as previously ROPEs
bf_rope_type_110 <- bayesfactor_rope(posterior = narrow_model,
                 null = rope,
                 parameter = 'b_distractor_type1')

bf_rope_related_110 <- bayesfactor_rope(posterior = narrow_model,
                 null = rope,
                 parameter = 'b_distractor_relatedness1')

bf_rope_inter_110 <- bayesfactor_rope(posterior = narrow_model,
                 null = rope,
                 parameter = 'b_distractor_type1:distractor_relatedness1')
bf_rope_type_110
bf_rope_related_110
bf_rope_inter_110
# our results
# type 0.0000245
# relatedness 0.0000420
# interaction 0.000182


# setting priors that are twice as wide as our original priors
priors_440 <- c(prior(normal(718, 220), class = "Intercept"),
                 prior(normal(0, 440), class = "b", coef = "distractor_relatedness1"),
              prior(normal(0, 440), class = "b", coef = "distractor_type1"),
              prior(normal(0, 440), class = "b", coef = "distractor_type1:distractor_relatedness1"),
              prior(lkj(1), class = 'cor'),
              # changing the parameter of the sd-priors from 3 to 300
              prior(cauchy(0, 300), class = 'sd', group = 'distractor_word'),
              prior(cauchy(0, 300), class = 'sd', group = 'participant'),
              prior(cauchy(0, 300), class = 'sd', group = 'pic'),
              prior(normal(0, 138), class = "sigma"))
# running model with new priors
wide_model <- brm(
  formula = model,
  data = aud,
  prior = priors_440,
  family = gaussian(),
  chains = 4,
  iter = 2000,
  warmup = 1000,
  thin = 1, 
  seed = 42,
  cores = 4        
)
# re-running our BF with same rope as previously ROPEs
bf_rope_type_440 <- bayesfactor_rope(posterior = wide_model,
                 null = rope,
                 parameter = 'b_distractor_type1')

bf_rope_related_440 <- bayesfactor_rope(posterior = wide_model,
                 null = rope,
                 parameter = 'b_distractor_relatedness1')

bf_rope_inter_440 <- bayesfactor_rope(posterior = wide_model,
                 null = rope,
                 parameter = 'b_distractor_type1:distractor_relatedness1')
bf_rope_type_440
bf_rope_related_440
bf_rope_inter_440
# our results
# type 0.0000120
# relatedness 0.0000157
# interaction 0.0000994
```

```{r, sensitivity analyses - smaller rope}
# rerunning the three different models with smaller rope
rope25 <- c(-25,25)
# As we dont have a very good reason, other than this number being the best we
# could find in the literature, foo setting our ropes at [-50, 50], we decided to
# look at how the BF is changing for the three model if our ROPE is just halt the 
# size.
# narrow priors model
bf_rope25_type_110 <- bayesfactor_rope(posterior = narrow_model,
                 null = rope25,
                 parameter = 'b_distractor_type1')

bf_rope25_related_110 <- bayesfactor_rope(posterior = narrow_model,
                 null = rope25,
                 parameter = 'b_distractor_relatedness1')

bf_rope25_inter_110 <- bayesfactor_rope(posterior = narrow_model,
                 null = rope25,
                 parameter = 'b_distractor_type1:distractor_relatedness1')
bf_rope25_type_110
bf_rope25_related_110
bf_rope25_inter_110
# our results - rope = 25, narrow priors model
# type 0.016
# relatedness 0.063
# interaction 0.014

# original model
bf_rope25_type <- bayesfactor_rope(posterior = proper_model,
                 null = rope25,
                 parameter = 'b_distractor_type1')

bf_rope25_related <- bayesfactor_rope(posterior = proper_model,
                 null = rope25,
                 parameter = 'b_distractor_relatedness1')

bf_rope25_inter <- bayesfactor_rope(posterior = proper_model,
                 null = rope25,
                 parameter = 'b_distractor_type1:distractor_relatedness1')
bf_rope25_type
bf_rope25_related
bf_rope25_inter
# our results - rope = 25, original model
# type 0.007
# relatedness 0.027
# interaction 0.008

# wide priors model
bf_rope25_type_440 <- bayesfactor_rope(posterior = wide_model,
                 null = rope25,
                 parameter = 'b_distractor_type1')

bf_rope25_related_440 <- bayesfactor_rope(posterior = wide_model,
                 null = rope25,
                 parameter = 'b_distractor_relatedness1')

bf_rope25_inter_440 <- bayesfactor_rope(posterior = wide_model,
                 null = rope25,
                 parameter = 'b_distractor_type1:distractor_relatedness1')
bf_rope25_type_440
bf_rope25_related_440
bf_rope25_inter_440
# our results - rope = 25, wide priors model
# type 
# relatedness 
# interaction 
```

```{r, sensitivity analyses - "in between" rope}
# rerunning the three different models with rope that's in between our original 
# [-50,50] and the previous [-25,25]
rope37.5 <- c(-37.5,37.5)
# For our third ROPE we decided to go with the middle between the large and small
# interval used before. 
# narrow priors model
bf_rope37.5_type_110 <- bayesfactor_rope(posterior = narrow_model,
                 null = rope37.5,
                 parameter = 'b_distractor_type1')

bf_rope37.5_related_110 <- bayesfactor_rope(posterior = narrow_model,
                 null = rope37.5,
                 parameter = 'b_distractor_relatedness1')

bf_rope37.5_inter_110 <- bayesfactor_rope(posterior = narrow_model,
                 null = rope37.5,
                 parameter = 'b_distractor_type1:distractor_relatedness1')
bf_rope37.5_type_110
bf_rope37.5_related_110
bf_rope37.5_inter_110
# our results - rope = 25, narrow priors model
# type 0.016
# relatedness 0.063
# interaction 0.014

# original model
bf_rope25_type <- bayesfactor_rope(posterior = proper_model,
                 null = rope25,
                 parameter = 'b_distractor_type1')

bf_rope25_related <- bayesfactor_rope(posterior = proper_model,
                 null = rope25,
                 parameter = 'b_distractor_relatedness1')

bf_rope25_inter <- bayesfactor_rope(posterior = proper_model,
                 null = rope25,
                 parameter = 'b_distractor_type1:distractor_relatedness1')
bf_rope25_type
bf_rope25_related
bf_rope25_inter
# our results - rope = 25, original model
# type 0.007
# relatedness 0.027
# interaction 0.008

# wide priors model
bf_rope25_type_440 <- bayesfactor_rope(posterior = wide_model,
                 null = rope25,
                 parameter = 'b_distractor_type1')

bf_rope25_related_440 <- bayesfactor_rope(posterior = wide_model,
                 null = rope25,
                 parameter = 'b_distractor_relatedness1')

bf_rope25_inter_440 <- bayesfactor_rope(posterior = wide_model,
                 null = rope25,
                 parameter = 'b_distractor_type1:distractor_relatedness1')
bf_rope25_type_440
bf_rope25_related_440
bf_rope25_inter_440
# our results - rope = 25, wide priors model
# type 
# relatedness 
# interaction 
```

## Interpreting the results

```{r, interpreting}
#| label: interpret_sim

summary(proper_model)

proper_model |> 
  gather_draws(b_distractor_relatedness1, b_distractor_type1, `b_distractor_type1:distractor_relatedness1`) |> 
  # plot
  ggplot(aes(x = .value,
             y = .variable)) +
    stat_halfeye() +
    geom_vline(xintercept = 0,
               lty = "dashed") +
  labs(x = "posterior estimate for reaction times",
       y = "") +
  theme_minimal()


p_direction(proper_model)


```
